### 2021-05-09 papers/SYRCoSE/intro.tex
Typos and grammar mostly

Under the interleaving semantics 
it has $3! = 6$ traces with each trace consisting of $4$ events,
as depicted in \cref{fig:intro-traces}.
Events themselves represent atomic side-effects produced by instruction executions.
In our case an event is either a \emph{write} of a value $a$ to a shared variable $x$ denoted as $\wlab{x}{a}$,
or a \emph{read} of a value $a$ from a shared variable $x$ denoted as $\rlab{x}{a}$.

The same information can be encoded in a single 
event structure containing $6$ events in total
(see \cref{fig:intro-es}). 
In the event structure there are two types of edges 
between the events. The grey arrows $e_1 \arrowCA e_2$ 
represent the \emph{causality relation}, a 
partial order reflecting the causal relationship
between the atomic events of computation.
The red edges $e_1 \arrowCF e_2$ represent 
the \emph{conflict relation} which is 
a symmetric and irreflexive relation 
encoding mutually exclusive events.
Each particular trace can be extracted from the event structure
as a linearisation of some \emph{configuration}, 
that is a causally-closed and conflict-free subset of events,  
which additionally should satisfy the constraint 
that each read is preceded by a matching write.

@@

In this work-in-progress report we sketch 
the common design principles behind our library
and give a concrete example of its usage  
by developing a formal mechanised semantics of simple 
register machine with shared memory.

>>>>

Under the interleaving semantics,
it has $3! = 6$ traces with each trace consisting of $4$ events,
as depicted in \cref{fig:intro-traces}.
Events themselves represent atomic side-effects produced by instruction executions.
In our case, an event is either a \emph{write} of a value $a$ to a shared variable $x$ denoted as $\wlab{x}{a}$,
or a \emph{read} of a value $a$ from a shared variable $x$ denoted as $\rlab{x}{a}$.

The same information can be encoded in a single 
event structure containing $6$ events in total
(see \cref{fig:intro-es}). 
In the event structure, there are two types of edges 
between the events. The grey arrows $e_1 \arrowCA e_2$ 
represent the \emph{causality relation}, a 
partial order reflecting the causal relationship
between the atomic events of computation.
The red edges $e_1 \arrowCF e_2$ represent 
the \emph{conflict relation} which is 
a symmetric and irreflexive relation 
encoding mutually exclusive events.
Each particular trace can be extracted from the event structure
as a linearisation of some \emph{configuration}, 
that is a causally-closed and conflict-free subset of events,  
which additionally should satisfy the constraint 
that each read is preceded by a matching write.

@@

In this work-in-progress report, we sketch 
the common design principles behind our library
and give a concrete example of its usage  
by developing a formal mechanised semantics of a simple 
register machine with shared memory.

### 2021-05-09 papers/SYRCoSE/overview.tex
Typos and grammar mostly

In this section we sketch the design principles of our library.

@@

The small-scale reflection approach is based on 
pervasive use of symbolic representations \emph{intermixed}
with logical ones within the confines of the same proof goal,
as opposed to large-scale reflection which does not allow such mixing.
Symbolic representations are connected to the corresponding logical ones
via user-defined \emph{reflect predicates}.
The symbolic representation can be manipulated 
by the computational engine of the language, 
allowing the user to automate low-level routine 
proof management by using various decision 
and simplification procedures.
Whenever the user needs to guide the proof 
it can switch to logical representation
and perform some proof steps manually.

To achieve better automation and e.g. get \emph{proof irrelevance}
automatically, one is encouraged
to use \emph{decidable} and \emph{computable} procedures
whenever possible.
For example, in the context of our library, 
we encode the binary relations of the event structures
as decidable \texttt{bool}-valued relations, 
\ie $\caOrd, \cfRel : E \fun E \fun \texttt{bool}$,
as opposed to \emph{propositional} 
relations of type $E \fun E \fun \texttt{Prop}$.

@@

We also favour the computational encoding of semantics. 
Similarly to the recent related works on mechanisation 
of operational semantics~\cite{Xia-al:POPL2019, Letan-al:CPP2020, Affeldt-al:ICMPC2019}, 
we encode the semantics as monadic interpreters.  
This allows us to extract~\cite{Letouzey:CCE2008} 
the semantics as a functional program and run it. 
We believe that the possibility to run the semantics 
is a very useful feature, as it allows 
to debug the formal semantics
and helps to develop better intuition about it.

To facilitate computable semantics, we define a subclass of
finitely supported event structures as a finite sequence of events
combined with finitely supported function which enhance events with 
additional information, such as their labels, causality predecessors, \etc 
Encoding finitely supported functions is not a trivial endeavor in
a proof assistant and for this task we use the \finmap~%
\footnote{\url{https://github.com/math-comp/finmap}}
library which is an extension of \mathcomp providing finite sets and
finite maps on types with a choice opeartor (rather than on finite types).

>>>>

In this section, we sketch the design principles of our library.

@@

The small-scale reflection approach is based on 
the pervasive use of symbolic representations \emph{intermixed}
with logical ones within the confines of the same proof goal,
as opposed to large-scale reflection which does not allow such mixing.
Symbolic representations are connected to the corresponding logical ones
via user-defined \emph{reflect predicates}.
The symbolic representation can be manipulated 
by the computational engine of the language, 
allowing the user to automate low-level routine 
proof management by using various decision 
and simplification procedures.
Whenever the user needs to guide the proof 
they can switch to the logical representation
and perform some proof steps manually.

To achieve better automation and e.g. get \emph{proof irrelevance}
automatically, one is encouraged
to use \emph{decision} procedures whenever possible.
For example, in the context of our library, 
we encode the binary relations of the event structures
as decidable \texttt{bool}-valued relations, 
\ie $\caOrd, \cfRel : E \fun E \fun \texttt{bool}$,
as opposed to \emph{propositional} 
relations of type $E \fun E \fun \texttt{Prop}$.

@@

We also favour the computational encoding of semantics. 
Similar to the recent related works on mechanisation 
of operational semantics~\cite{Xia-al:POPL2019, Letan-al:CPP2020, Affeldt-al:ICMPC2019}, 
we encode the semantics as monadic interpreters.  
This allows us to extract~\cite{Letouzey:CCE2008} 
the semantics as a functional program and run it. 
We believe that the possibility to run the semantics 
is a very useful feature, as it allows 
to debug the formal semantics
and helps to develop better intuition about it.

To facilitate computable semantics, we define a subclass of
finitely supported event structures as a finite sequence of events
combined with finitely supported function which enhances events with 
additional information, such as their labels, causality predecessors, \etc 
Encoding finitely supported functions is not a trivial endeavor in
a proof assistant and for this task we use the \finmap~%
\footnote{\url{https://github.com/math-comp/finmap}}
library which is an extension of \mathcomp providing finite sets and
finite maps on types with a choice operator (rather than on finite types).

### 2021-05-09 papers/SYRCoSE/regmachine.tex
Use xspace to handle spaces after \newcommands such as \coq
Typos and grammar mostly

For our case study we use a simple idealised model 
of a register machine, which consists of a finite 
sequence of instructions, an instruction pointer 
and an infinite set of registers. 
The syntax of the machine's language is 
shown in~\cref{fig:regmachine-syntax}.

@@

The semantics is given in the form of a
\emph{labelled transition system}:~%
\footnote{As we have mentioned, in our \coq~development we actually use 
the monadic encoding of the operational semantics. 
The labelled transition system can be derived from this encoding.}
$P \vdash s \thrdstep{l} s'$,
where $P$ is a program, $l$ is a label, $s$ and $s'$ are states of the machine. 
The state of the machine itself consist of an instruction pointer $\ip{i}$
and a map from registers to their values $\regmap$, 
as shown in \cref{fig:regmachine-thrdstate}. 
The rules of the semantics are standard 
(see \cref{fig:regmachine-thrdsem}).

>>>>

For our case study, we use a simple idealised model 
of a register machine, which consists of a finite 
sequence of instructions, an instruction pointer,
and an infinite set of registers. 
The syntax of the machine's language is 
shown in~\cref{fig:regmachine-syntax}.

@@

The semantics is given in the form of a
\emph{labelled transition system}:~%
\footnote{As we have mentioned, in our \coq development we actually use 
the monadic encoding of the operational semantics. 
The labelled transition system can be derived from this encoding.}
$P \vdash s \thrdstep{l} s'$,
where $P$ is a program, $l$ is a label, $s$ and $s'$ are states of the machine. 
The state of the machine itself consists of an instruction pointer $\ip{i}$
and a map from registers to their values $\regmap$, 
as shown in \cref{fig:regmachine-thrdstate}. 
The rules of the semantics are standard 
(see \cref{fig:regmachine-thrdsem}).

### 2021-05-09 papers/SYRCoSE/related.tex
Typos and grammar mostly

More recently event structures were applied 
in the context of relaxed memory models~%
\cite{Jeffrey-Riely:LICS16, PichonPharabod-Sewell:POPL16, Chakraborty-Vafeiadis:POPL19, Moiseenko-al:ECOOP20}.
Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
that was accompanied with a mechanisation in a proof assistant. 
The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
memory model in \coq. However, this memory model uses 
custom variant of event structures, that does not 
obey axioms of any conventional class of 
event structures~\cite{Winskel:82, Langerak:91, Boudol-Castellani:1991}. 
This fact makes it harder to reuse and adapt it 
to other applications of the theory.

>>>>

More recently event structures were applied 
in the context of relaxed memory models~%
\cite{Jeffrey-Riely:LICS16, PichonPharabod-Sewell:POPL16, Chakraborty-Vafeiadis:POPL19, Moiseenko-al:ECOOP20}.
Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
that was accompanied by a mechanisation in a proof assistant. 
The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
memory model in \coq. However, this memory model uses 
a custom variant of event structures, that does not 
obey axioms of any conventional class of 
event structures~\cite{Winskel:82, Langerak:91, Boudol-Castellani:1991}. 
This fact makes it harder to reuse and adapt it 
to other applications of the theory.
